Measurement of the positive kaon-argon total hadronic differential cross section in the LArIAT experiment

### 2018-03-16 Chapter-2/2-TheLArIATExperiment.tex
s

\subsubsection{Multi-Wire Proportional Chambers functionality}\label{sec:MWPCfunc}

@@

\section{In the Cryostat}
\subsection{LArTPC: Charge Collection}\label{sec:TPCCharge}
The LArIAT Liquid Argon Time Projection Chamber is a rectangular box of dimensions 47 cm (width) x 40 cm (height) x 90 cm (length), containing 170 liters of Liquid Argon.
The LArTPC three major subcomponents are 
\begin{itemize} 
\item[1)] the cathode and field cage,
\item [2)] the wire planes, 
\item [3)] the read-out electronics. %
\end{itemize}

@@

\subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.

%%%%%%%%%%%%%%%%%%%%%
As in any other LArTPC, argon purity is a crucial parameter for LArIAT. Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt). We achieve this level of purity in several stages. The specifics required for the commercial argon bought for LArIAT are 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen. This argon is monitored with the use of commercial gas analyzer.
Argon is stored in a dewar external to LArIAT hall and filtered before filling the TPC. %The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.  The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.  The piping was cleaned to remove oil and grease before being welded into the system. 
LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) apt to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}. A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated. In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s}.
The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat. Argon is not recirculated in the system, but it rather boils off and vent to the atmosphere. During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged. In fact, we need to monitor the level, temperatures, and pressures  of the argon both in the commercial dewar and inside the cryostat.

\section{Trigger and DAQ}
%The read-out of a LArIAT event is triggerred by a CAEN V1495 module.

%Scope of the LArIAT DAQ and trigger system is to start the read out of the detector, to convert signals from analog to digital, to interface and synchronize with the accelerator system, to store raw data on disk, transfer data to archival storage and conduct a first pass data quality monitoring.

%A CAEN V1495 governs LArIAT DAQ system.

%The system is made by commercial VME electronics from CAEN, a VME time to digital converter for the beamline time of flight system, and a readout system for beamline wire chambers.

>>>>

\subsubsection{Multi-Wire Proportional Chambers functionality}\label{sec:MWPCfunc}
We use the wire chamber system together with the bending magnets to measure the particles momentum.

In the simplest scenario, only one hit on each and every of the four wire chambers is recorder during a single readout of the detector systems.  Thus, we use the hits position in the two wire chambers upstream to the magnets to form a trajectory before the bent and the hits position in the two wire chambers downstream to the magnets to form a trajectory after the bent. We use the angles in the XZ plane between the upstream and downstream trajectories  to calculate the z component of the momentum as follows:

\begin{equation}
P_z=\frac{B_{eff}L_{eff}}{3.3(sin(\theta_{DS})-sin(\theta_{US}))},
\label{eq:momformula}
\end{equation}

where $B_{eff}$ is the effective maximum field in a square field approximation,  $L_{eff}$ is the effective length of both magnets (twice the effective length of one magnet) and  3.3~$c^{-1}$ the conversion factor from from T*m to MeV/c. By using the hit positions on WC3 and WC4, we estimate the azimutal and polar angles of the particle trajectory, and we are able to calculate the other components of the momentum.

This simple scenario can get complicated in two ways:  the presence of multiple hits in a single wire chamber or by the absence of hits in one (or more) wire chamber. The first complication is due to beam pile up, while the latter is due to wire chamber inefficiency. Even in the case of multiple hits, at most one wire chamber track is reconstructed per event; since the magnets bend particles only in the X direction, we assume the particle trajectory to be roughly constant in the YZ plane, thus we keep the combination of hits most coplanar on YZ plane.  It is possible to reconstruct the particle momentum even in case of wire chamber inefficiency, if the information is missing in either of the two middle wire chambers (WC2 or WC3). %Assuming identical magnets, the upstream and downstream trajectories projected without bending should intersect at some point in the space between the magnets. Over a collection of many tracks, these points of intersection would trace a plane, therefore the term "midplane" is used. The completed half of the wire chamber track is projected to the midplane and the point of intersection is used with the incomplete half to finish the track.

\textcolor{red}{BETTER EXPLAINATION OF HIGH YIELD}
\textcolor{red}{UNCERTAINTY ON MOMENTUM MEASUREMENT}
Events satisfying the simplest scenario of one single hit in each of the four wire chambers form the so called ``Picky track" sample.  We construct an other, higher statistics sample where we loose the requirements on single hit and wire chamber efficiency called "High Yield". For LArIAT Run II, the "High Yield" sample is 
\textcolor{red}{ HIGH YIELD increase in stat} the "Picky Tracks" sample. We use Picky tracks for calibration purposes: \textcolor{red}{ I DONT UNDERSTAND THIS Plotting the fractional error of the corrected three-point momentum compared to the known four-point momentum shows an uncertainty of approximately 2\%.}

@@

\section{In the Cryostat}
\subsection{Cryogenics and Argon Purity}\label{ch:Cryo}
LArIAT repurposed the ArgoNeuT cryostat \cite{argoneut} in order to use it in a beam of charge particles. We also added a new process piping and a new liquid argon filtration system in FTBF.
Inside the LArIAT experimental hall, the cryostat sits on the beam of charge particles with its horizontal main axis oriented parallel to the beam.

%%%%%%%%%%%%%%%%%%%%%
As in any other LArTPC, argon purity is a crucial parameter for LArIAT. Indeed, the presence of contaminants effects both the basic working principles of a LArTPC: electronegative contaminants such as oxygen and water decrease the number of ionization electrons collected on the wires after drifting through the volume, while contaminants such as Nitrogen decrease the light yield from scintillation light, especially in its slow component.
In LArIAT, contaminations should not exceed the level of 100 parts per trillion (ppt). We achieve this level of purity in several stages. The specifics required for the commercial argon bought for LArIAT are 2 parts per million (ppm) oxygen, 3.5~ppm water, and 10~ppm nitrogen. This argon is monitored with the use of commercial gas analyzer.
Argon is stored in a dewar external to LArIAT hall and filtered before filling the TPC. %The argon is delivered from the commercial dewar to the cryostat through 2.54~cm diameter schedule 10 stainless steel piping.  The piping was insulated with 20.32~cm of polyurethane foam by the manufacturer.  The piping was cleaned to remove oil and grease before being welded into the system. 
LArIAT uses a filtration system designed for the Liquid Argon Purity Demonstrator (LAPD)~\cite{LAPD}: half of a 77~liter filter contains a 4A molecular sieve (Sigma-Aldrich~\cite{sigma-aldrich}) apt to remove mainly water, while the other half contains BASF~CU-0226~S, a highly dispersed copper oxide impregnated on a high surface area alumina, apt to remove mainly oxygen~\cite{basf}. A single pass of argon in the filter is sufficient to achieve the necessary purity, unless the filter is saturated. In case the filter saturates, the media needs to be regenerated by using heated gas; this happened twice during the Run II period\footnote{We deemed the filter regeneration necessary every time the electron lifetime dropped under 100 \textmu s}.
The filtered argon reaches the inner vessel via a liquid feedthrough on the top of the cryostat. Argon is not recirculated in the system, but it rather boils off and vent to the atmosphere. During data taking, we replenish the argon in the cryostat several times per day to keep the TPC high voltage feedthrough and cold electronics always submerged. In fact, we need to monitor the level, temperatures, and pressures  of the argon both in the commercial dewar and inside the cryostat. 
\subsection{LArTPC: Charge Collection}\label{sec:TPCCharge}
The LArIAT Liquid Argon Time Projection Chamber is a rectangular box of dimensions 47 cm (width) x 40 cm (height) x 90 cm (length), containing 170 liters of Liquid Argon.
The LArTPC three major subcomponents are 
\begin{itemize} 
\item[1)] the cathode and field cage,
\item [2)] the wire planes, 
\item [3)] the read-out electronics. %
\end{itemize}

@@

\section{Trigger and DAQ}
The LArIAT DAQ and trigger system governs the read out of all the many detectors forming LArIAT. 
The CAEN V1495 module featuring a user-programmable FPGA  is the core of this system.  Every 10~ns, this module checks for matches between sixteen logical inputs and user-defined patterns in the trigger menu; if it finds a match for two consecutive clock ticks, that trigger fires.

% converts signals from analog to digital, interfaces and synchronizes with the accelerator system, stores raw data on disk, transfers data to archival storage and conduct a first pass data quality monitoring.

### 2018-03-16 Chapter-4/3-KaonInteractionsInLAr.tex
s

>>>>

\subsection{Uncertainty budget}
Measuring an hadronic cross section  in LArIAT translates into counting how many hadrons impinged on a slab of argon at a given energy and how many of those hadrons interacted at said energy. So, the key questions here are:
\begin{itemize}
\item[]a) how well do we know the kinetic energy at each point of the tracking? %(Incident Kinetic Energy bins)
\item[]b) how well do we know when the tracking stops? %(Interacting Kinetic Energy bin)
\item[]c) are there any systematic shifts?
\end{itemize}

In order to answer this question, will discuss first a simple scenario  were our beam is 100\% made of pions which arrive as primaries in the TPC (no decay in the beam and no inelastic interaction before the TPC front face). We will then add a layer of complexity by discussing how we handle beamline contamination.
\subsubsection{Pure beam of pions}
\subsubsection{Handling beamline contamination}
What is the beamline contamination? We define beamline contamination every TPC track matched to the WC track which is not a primary pion. There are 4 different types of beamline contaminations:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondary pions,
\item[]4) matched pile up events.
\end{itemize}
So, how do we handle this contamination?
The first step is understanding what percentage of events used in the data cross section calculation is not a primary pion.

The number of electrons and muons in the pion sample is estimated via the beamline MC. We simulate the beamline composition for each of the magnet settings present on data (-60A and -100A). 
and we weight for the number of events which pass the mass cut in both the configurations. 
For each beamline

We estimate the matched pile up events to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match\footnote{ Events with multiple matches are rejected.} with a pile up event in the absence of a real particle  within a four centimeter radius from the WC track projection extremely small.
